---
title: "Analysis of DENV in Brazil: local incidence"
output:
  html_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par(mgp = c(2, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin3 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", mai = rep(.1, 4))
    else NULL
  }
)

knitr::opts_chunk$set(echo       = TRUE,
                      cache      = TRUE, 
                      margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center")
```

```{css, echo=FALSE}
pre {
max-height: 300px;
overflow-y: auto;
}

pre[class] {
max-height: 100px;
}
```

## Links to local data sets

Here we use the following links to data.

```{r}
# directories
CASE_DATA_FOLDER <- "data/bra2_cases/"
SF_DATA_FOLDER <- "data/bra2_sf/"
POPULATION_DATA_FOLDER <- "data/bra2_population/"
SUMMARY_OUTPUT_DATA_FOLDER <- "output/data/summary_output/"

# raw data 
bra1_path <- file.path(SF_DATA_FOLDER, "bra1_sf_data.rds")
bra2_path <- file.path(SF_DATA_FOLDER, "simplified_bra2_sf_data.rds")
bra1_macroregions_path <- file.path(SF_DATA_FOLDER, "bra1_macroregions.xlsx")
bra2_pop_data_path <- file.path(POPULATION_DATA_FOLDER, "bra2_population_data_2000_2014.rds")
bra2_case_data_path <- file.path(CASE_DATA_FOLDER, "bra_DENV_case_data_2000_2014.rds")

# summary data
bra2_log_incidence_typical_year_summary_data_path <- file.path(SUMMARY_OUTPUT_DATA_FOLDER,
                                                               "region_typical_year_log_incidence_data.Rdata")
```

Here, we define the file paths to the output. 

```{r}
SUMMARY_PLOTS_FOLDER <- "output/plots/"
SUMMARY_RESULTS_FOLDER <- "output/results/"
```

Change them accordingly if you want to run the script locally on your computer.

## Preambule

This document focuses on describing local DENV incidence across space (municipality level) and across time (2000-2014). 

## Packages

The needed packages:

```{r message = FALSE}
library(pacman)
p_load(tidyverse, sf, RColorBrewer, readxl, magrittr, ggh4x)
```

## Raw data

### Case data

We read in the raw case data at the municipality level in Brazil. 

```{r}
bra2_case_data <- readRDS(file=bra2_case_data_path)
```

Let's get the state/city to city id mapping, so we can add it to other sets of data for ease of data management later on. 

```{r}
city_cityid_mapping <- bra2_case_data %>% 
  select(city_id, state, city) %>% 
  unique()
```

### Spatial data 

Here, we read in and format the spatial polygons data for Brazil. 

```{r}
bra1_macroregions <- read_excel(path=bra1_macroregions_path)

bra1 <- readRDS(bra1_path) %>% 
  mutate(x=st_coordinates(st_centroid(geometry))[,1], y=st_coordinates(st_centroid(geometry))[,2]) %>% 
  left_join(bra1_macroregions, by="state")

bra2 <- readRDS(bra2_path) %>%
  left_join(city_cityid_mapping, by=c("state", "city"))
```

### Population data

Here, we read in the population data at the municipality level in Brazil. 

```{r}
bra2_pop_data <- readRDS(file=bra2_pop_data_path) %>% 
  left_join(city_cityid_mapping, by=c("state", "city"))
```

Let's also add population density data at the municipality level, which will be used extensively in the following analyses. 

```{r}
bra2_pop_data %<>%
  left_join(st_drop_geometry(bra2), by=c("state", "city", "city_id")) %>% 
  mutate(den_km2=n/area_m2*10^6)
```

## Processed summary data 

### Case data

```{r}
load(bra2_log_incidence_typical_year_summary_data_path) #region_typical_year_cases_data
bra2_log_incidence_typical_year_summary_data <- region_typical_year_cases_data
rm(region_typical_year_cases_data)
```

## DENV incidence

In this section, we simply describe DENV incidence across Brazil. 

### Data management

There are four municipalities with incomplete case data for the period from 2000 to 2014: Pará, Mojuí dos Campos; Santa Catarina, Pescaria Brava; Santa Catarina, Balneário Rincão; Mato Grosso do Sul, Paraíso das Águas. 

+ Santa Catarina, Pescaria Brava was established in 2013. 
+ Balneário Rincão; Mato Grosso do Sul was established in 2013. 
+ Pará, Mojuí dos Campos was established in 2013. 
+ Mato Grosso do Sul, Paraíso das Águas has incomplete data for all years except 2004 and 2014. 

Let's consolidate the data for ease of presentation. 

```{r}
# compute the mean population in each municipality over the entire period
summarised_bra2_pop_data <- bra2_pop_data %>%
  group_by(state, city, city_id) %>% 
  summarise(n=mean(n, na.rm=TRUE), .groups="keep")

# compute incidence in each region
incidence_data <- bra2_case_data %>% 
  filter(!grepl(pattern="Município ignorado", x=city, fixed=TRUE)) %>% 
  group_by(city_id, state, city, year) %>%
  summarise(cases=sum(cases, na.rm=FALSE)+1, .groups="keep") %>%
  left_join(bra2_pop_data, by=c("state", "city", "city_id", "year")) %>%
  mutate(incidence=cases/n) %>% 
  ungroup() %>%
  group_by(city_id, state, city) %>%
  summarise(mean_yearly_incidence=mean(incidence, na.rm=FALSE)*10^5, 
            cum_cases=sum(cases, na.rm=FALSE), .groups="keep") %>%
  as.data.frame() %>%
  left_join(summarised_bra2_pop_data, by=c("state", "city", "city_id")) %>%
  mutate(cum_incidence=cum_cases/n*10^5)

aggregated_mun_monthly_case_data <-  bra2_case_data %>% 
  left_join(bra2_pop_data, by=c("state", "city", "city_id", "year")) %>% 
  group_by(city_id, state, city, year, month) %>% 
  summarise(cases=sum(cases, na.rm=FALSE), n=n, .groups="keep") %>%
  as.data.frame() %>% 
  mutate(date=as.Date(paste0(year, "-", month, "-", "2"))) %>% 
  mutate(incidence=cases/n*10^5)

state_cum_incidence_data <- bra2_case_data %>% 
  left_join(select(st_drop_geometry(bra1), state, region), by="state") %>% 
  left_join(bra2_pop_data, by=c("state", "city", "city_id", "year")) %>% 
  group_by(region, state, year, month) %>% 
  summarise(cases=sum(cases, na.rm=TRUE), 
            n=sum(n, na.rm=TRUE), .groups="keep") %>% 
  ungroup() %>% group_by(region, state) %>% 
  summarise(cum_incidence=sum(cases)/mean(n)*10^5, .groups="keep") %>% 
  as.data.frame()
```

### Figures

Here, we plot a summary figure for the local DENV incidence across Brazil. 

```{r}
plot_DENV_incidence_plot <- function(incidence_data, state_cum_incidence_data, aggregated_mun_monthly_case_data, 
                                     bra2_incidence_typical_year_summary_data, 
                                     bra2, bra1, file_name) {
  # necessary colour palette
  pal <- rev(brewer.pal(11, "Spectral"))
  full_pal2 <- c(colorRampPalette(pal[1:3])(8), pal[4:7], colorRampPalette(pal[8:11])(8))
  
  # organise incidence data
  incidence_data %<>% 
    mutate(log10_mean_yearly_incidence = log10(mean_yearly_incidence), 
           log10_cum_cases=log10(cum_cases+1)) %>% 
    left_join(bra2, by=c("state", "city", "city_id")) %>% 
    st_as_sf()
  
  # map of mean yearly incidence in each muncipality
  focal_cities <- c(520870, 150140, 320530, 230440)
  x_adj <- c(-1, 1.1, 2.65, 1.5)
  y_adj <- c(-0.85, 1.8, 0, 1.5)
  label_adj <- data.frame(city_id=focal_cities, x_adj=x_adj, y_adj=y_adj)
  focal_bra2 <- bra2 %>% 
    filter(city_id %in% focal_cities) %>% 
    mutate(x=st_coordinates(st_centroid(geometry))[,1], y=st_coordinates(st_centroid(geometry))[,2]) %>%
    left_join(mutate(label_adj, city_id=as.character(city_id)), by="city_id")
  
  case_burden_plot <- ggplot() + 
    geom_sf(data=incidence_data, 
            mapping=aes(geometry=geometry, fill=log10_mean_yearly_incidence, color=log10_mean_yearly_incidence), lwd=0) + 
    geom_sf(data=bra1, 
            mapping=aes(geometry=geometry), color="black", fill=NA, lwd=0.1) + 
    geom_point(data=focal_bra2, mapping=aes(x=x, y=y), shape=24, fill="black", colour="black", size=1.5) + 
    geom_text(data=focal_bra2, mapping=aes(x=x+x_adj, y=y+y_adj, label=city), size=2) + 
    scale_fill_gradientn(colours=full_pal2, na.value="grey", 
                         breaks=seq(0, 3, 1), limits=c(-0.16, 3.5), labels=c("1", "10", "100", "1000"), name="Mean yearly\nincidence") +
    scale_color_gradientn(colours=full_pal2, na.value="darkgrey", guide="none", 
                          breaks=seq(0, 3, 1), limits=c(-0.16, 3.5), name=NULL) + 
    guides(fill=guide_colorbar(ticks.colour="black", frame.colour="black")) + 
    scale_x_continuous(limits=c(-73, -36)) +
    theme(panel.grid=element_blank(), axis.line=element_blank(),
          axis.text=element_blank(), axis.title=element_blank(),
          panel.background=element_blank(), axis.ticks=element_blank()) + 
    # theme(legend.position="right", legend.title.align=0, legend.justification="center", 
    #       legend.key.width=unit(0.3, "cm"), legend.title=element_text(size=8), legend.text=element_text(size=8), 
    #       legend.key.height=unit(0.5, "cm")) + 
    theme(legend.position=c(0.2, 0.28), 
          legend.key.size = unit(0.4, "cm"), legend.text=element_text(size=8), 
          legend.title=element_text(size=8), 
          legend.box.background=element_blank(), legend.background=element_blank()) + 
    theme(plot.margin=unit(c(0, 0, 0, 0), "cm"))
  
  # map of Brazil with macroregions and states labelled
  region_colors <- brewer.pal(5, "Set1")
  size_scale=0.5
  mod_bra1 <- bra1
  mod_bra1 %<>% 
    mutate(state=ifelse(state=="Distrito Federal", "Distrito\nFederal", state)) %>% 
    mutate(state=ifelse(state=="Mato Grosso do Sul", "Mato Grosso\ndo Sul", state)) %>% 
    mutate(state=ifelse(state=="Rio Grande do Norte", "Rio Grande\ndo Norte", state)) %>% 
    mutate(state=ifelse(state=="Rio Grande do Sul", "Rio Grande\ndo Sul", state))
  state_names <- sort(pull(st_drop_geometry(mod_bra1), state))
  x_adj <- c(-0.7, 3.2, 0, 0, 0, 0.2, 0, 4.5, 0, 0, 0, 0, 0.3, 0, 4.0, 0, 6.2, 0, 5.2, 3.7, 0, 0, 0.5, 5.5, -0.75, 2.4, 0)
  y_adj <- c(0.5, 0, 0.2, 0, 0, 0.5, 1.2, 0, -0.5, 0, 0, 0, -0.5, 0, 0, 0, 0, -0.5, 0, 1.5, 0, -0.6, -1, 0, 0.5, -0.8, -.7)
  size <- c(3, 3, 3, 3, 3, 3, 2, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3)*size_scale
  color <- c("white", region_colors[2], "white", "white", "white", "white", "white", 
             region_colors[4], "white", "white", "white", "white", "white", 
             "white", region_colors[2], "white", region_colors[2], "white", region_colors[4], 
             region_colors[2], "white", "white", "white", region_colors[5], "white", region_colors[2], "white")
  mod_bra1 %<>% 
    left_join(data.frame(state=state_names, x_adj=x_adj, y_adj=y_adj, size=size, color=color), by="state")
  legend_title_features <- data.frame(region=c("North", "Northeast", "Central-West", "Southeast", "South"), 
                                      color=region_colors, 
                                      x=c(-70, -35, -65, -40, -62), 
                                      y=c(4, 0, -20, -25, -30))
  state_plot <- ggplot(data=mod_bra1) + 
    geom_sf(mapping=aes(geometry=geometry, fill=region), color="white", lwd=0.1) + 
    geom_text(mapping=aes(x=x+x_adj, y=y+y_adj, label=state), size=size, color=color, fontface="bold") + 
    scale_fill_manual(breaks=c("North", "Northeast", "Central-West", "Southeast", "South"), 
                      values=brewer.pal(5, "Set1")) + 
    scale_x_continuous(limits=c(-73, -29)) + 
    theme(legend.title=element_blank(), legend.position=c(0.2, 0.3), 
          legend.key.size = unit(0.3, "cm"),
          legend.box.background=element_blank(), legend.background=element_blank()) + 
    theme(panel.grid=element_blank(), axis.line=element_blank(),
          axis.text=element_blank(), axis.title=element_blank(),
          panel.background=element_blank(), axis.ticks=element_blank()) + 
    theme(plot.margin=margin(0, 0, 0, 0))
  
  # barplot of cumulative incidence in each state for 2000-2014
  states <- state_cum_incidence_data %>% 
    arrange(desc(region), desc(state)) %>% 
    pull(state) %>% 
    unique()
  state_cum_incidence_data %<>%
    mutate(state=factor(state, states))
  state_cum_incidence_plot <- ggplot(data=state_cum_incidence_data) + 
    geom_bar(mapping=aes(x=state, y=cum_incidence, fill=region), color="black", stat="identity") + 
    scale_fill_manual(breaks=c("North", "Northeast", "Central-West", "Southeast", "South"), 
                      values=brewer.pal(5, "Set1")) + 
    labs(y="Cumulative incidence") + 
    theme_bw() + 
    theme(axis.title.y=element_blank(), axis.text=element_text(size=8)) + 
    theme(legend.position="none", panel.grid=element_blank(), axis.title=element_text(size=8)) + 
    coord_flip()
  
  # line plots of monthly incidence in four selected municipalities
  focal_cities <- c(520870, 150140, 320530, 230440)
  aggregated_mun_monthly_case_data %<>% 
    filter(city_id %in% focal_cities) %>% 
    mutate(region_name=paste0(city, ", ", state)) %>% 
    mutate(log10_incidence=ifelse(incidence<0.1, log10(0.1), log10(incidence)))
  number_months <- 12*15
  dates <- unique(pull(aggregated_mun_monthly_case_data, date))
  tick_labels <- rep("", number_months)
  tick_labels[seq(1, number_months, 12)] <- 2000:2014
  monthly_incidence_lineplot <- ggplot(data=aggregated_mun_monthly_case_data)  +
    geom_line(mapping=aes(y=log10_incidence, x=date, color=city_id)) +
    geom_ribbon(mapping=aes(x=date, ymin=-1, ymax=log10_incidence, fill=city_id), alpha=0.3) + 
    labs(y="Incidence") +
    scale_x_continuous(minor_breaks=dates,
                       breaks=dates[seq(1, number_months, 12)],
                       labels=paste("", 2000:2014, sep=""),
                       expand=c(0.01, 0.01), guide="axis_minor", name=NULL) +
    scale_fill_manual(breaks=focal_cities, values=brewer.pal(5, "Set1")[c(3, 1, 4, 2)]) + 
    scale_color_manual(breaks=focal_cities, values=brewer.pal(5, "Set1")[c(3, 1, 4, 2)]) + 
    scale_y_continuous(breaks=c(-1, 0, 1, 2, 3), labels=c("0.1", "1", "10", "100", "1000")) + 
    theme_bw() +
    theme(ggh4x.axis.ticks.length.minor=rel(0.5), axis.text.x=element_text(angle=90, vjust=0.6)) +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="none", 
          axis.text=element_text(size=6), axis.title=element_text(size=8), strip.text=element_text(size=8)) + 
    facet_wrap(~region_name)
  
  # line plot of DENV incidence during a typical year for four selected municipalities
  bra2_incidence_typical_year_summary_data %<>%
    filter(city_id %in% focal_cities) %>%
    mutate(city_id=factor(city_id, c(520870, 320530, 230440, 150140))) %>%
    mutate(cases=log10(exp(cases)-1)) %>% 
    mutate(month=factor(month, levels=c(7:12, 1:6)))
  
  typical_year_incidence_lineplot <- ggplot(data=bra2_incidence_typical_year_summary_data)  +
    geom_line(mapping=aes(y=cases, x=month, color=city_id, group=city_id)) +
    geom_ribbon(mapping=aes(x=month, ymin=0, ymax=cases, group=city_id, fill=city_id), alpha=0.5) + 
    labs(y="Incidence", x="Month") +
    scale_fill_manual(breaks=focal_cities, values=brewer.pal(5, "Set1")[c(3, 1, 4, 2)]) +
    scale_color_manual(breaks=focal_cities, values=brewer.pal(5, "Set1")[c(3, 1, 4, 2)]) +
    scale_y_continuous(limits=c(0, 2.11), breaks=c(0, 1, 2), labels=as.character(c(1, 10, 100)))  + 
    theme_bw() +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="none", 
          axis.text=element_text(size=8), axis.title=element_text(size=8))
  
  # combine plots into a single figure
  pAB <- cowplot::plot_grid(state_plot, case_burden_plot, labels=c("A", "B"), nrow=1)
  pD <- cowplot::plot_grid(state_cum_incidence_plot, typical_year_incidence_lineplot, labels=c("C", "E"), ncol=1, rel_heights=c(1, 0.5))
  pABD <- cowplot::plot_grid(pAB, monthly_incidence_lineplot, labels=c("", "D"), ncol=1)
  pABCD <- cowplot::plot_grid(pABD, pD, rel_widths=c(1, 0.5), nrow=1)
  scale_factor <- 0.45
  pdf(file.path(SUMMARY_PLOTS_FOLDER, file_name), w=8.27, h=11.69*scale_factor)
  print(pABCD)
  a <- dev.off()
}
plot_DENV_incidence_plot(incidence_data=incidence_data, 
                         aggregated_mun_monthly_case_data=aggregated_mun_monthly_case_data, 
                         state_cum_incidence_data=state_cum_incidence_data, 
                         bra2_incidence_typical_year_summary_data=bra2_log_incidence_typical_year_summary_data, 
                         bra2=bra2, bra1=bra1,
                         file_name="local_incidence_figure.pdf")
```